The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
Final round of correction based on Simon/Dave comments.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1857 4525493e-7705-40b1-a816-d608a930855b

\subsection{DAS Mapping DB\label{MappingDB}}
The Mapping DB is used to track services known to DAS and how to interact with them.
%The Mapping DB represents DAS {\it registration} service. 
It collects
information about data-service APIs, translates API input/output
parameters into DAS keys and contains mapping between DAS records
and their UI representation.

@@

\subsection{DAS caching system}
The DAS caching system is used to dynamically fetch
and aggregate data upon user requests. We use the following 
algorithm (code snippets are written in Python \cite{Python}):
\begin{enumerate}[1.]
\item Decompose query into set of conditions (\verb+conds+) 
and selection keys (\verb+skeys+).
\item Check cache for query hash. If found, look-up results from the cache,
otherwise proceed to the next step.
\item Identify list of APIs which can serve requested data
\begin{verbatim}
apilist = [r for r in mapping_db(skeys, conds)]
\end{verbatim}
Here \verb+r+ represents a triplet \verb+url, api, args+ and 
\verb+mappind_db+ defines translation between input query and data-service
APIs which can delivery the data.
\item Iterate over APIs
\begin{verbatim}
for url, api, args in apilist:
    if  superset_in_analytics(api, args):
        continue
    for row in parser(get_data(url, api, args)):
        yield row
\end{verbatim}
Here \verb+superset_in_analytics+ checks if provided \verb+api, args+
are subset of already registered in Analytics DB api call(s). The 
\verb+get_data+ and \verb+parser+ are define methods to fetch and parse
data, respectively.
\item Perform aggregation by iterating over records from the previous step
\begin{verbatim}
if cache_has(daskey):
   existing_record = get_from_cache(daskey)
   merge(row, existing_record)
   insert_into_cache(row)
   delete_from_cache(existing_record)
else:
   insert_into_cache(row)
\end{verbatim}
The \verb+daskey+
is a key which identifies the record defined in Mapping DB, e.g. \verb+file.name+
for \verb+file+ record.
We did use a bulk insert operations at this step to increase throughput. 
\item Get results from the cache
\end{enumerate}
We also have an option to run independent set of daemons, DAS robots, who
monitor Analytics DB for most popular requests and pre-fetch the most
common data into the cache.

\subsection{DAS queries}
All DAS queries are expressed in a free text-based form, either as a 
set of keywords or key-value pairs, where a pair can represent
condition, e.g. \verb+key > value+. We use a pre-defined set of keys registered 
in DAS to describe common entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
is based on the analysis of daily work operations of our users, see \cite{DBS07}.
Each key represents a DAS record. Every record is stored into the cache
in JSON format. Due to the ``schemaless'' nature of underlying back-end (MongoDB)
we are able to store DAS records of arbitrary structure, e.g.
dictionary, lists, key-value pairs, etc. Therefore every DAS key has a 
set of attributes representing JSON structure. For instance a document
\begin{verbatim}
{"block": {"name": "abc", "size": 1, "replica": [...]}}
\end{verbatim}
is associated with a \verb+block+ key who has the following attributes
\verb+name+, \verb+size+, \verb+replica+. The set of attributes is based 
on a set of meta-data information stored in a record and therefore
a nested structures are allowed, e.g. \verb+block.replica.size+. Our users 
are able to specify a context of their queries by using appropriate 
attribute(s). For example, query \verb+block+ represents search for all
block records rather then ``block'' string, while query
\verb+block.name=abc*+ instructs the system to look-up appropriate blocks
with matched names. The numerical values, conditions, grouping are
easy to express too, e.g. \verb+run > 10 and run < 100+ query 
is used to get all run records for provided range of run numbers. 
To complement the strength of the queries, we do plan to index 
dynamically DAS records and add a keyword-based 
search within record content later.

All input queries are transformed into MongoDB query syntax who is itself
a JSON record. It contains a set of selection 
keys (\verb+fields+) and conditions (\verb+spec+). For example, a query
\verb+dataset, file=abc, run>10+
is represented as
\begin{verbatim}
{"fields": ["dataset"], "spec": {"file": "abc", "run": {"$gt": 10}}}
\end{verbatim}
Therefore users queries are stored into Analytics DB similar to other DAS 
records and used for further analysis.

\subsection{Data-services and aggregation}
As was shown in section \ref{MappingDB} the DAS Mapping DB became an authorative
source of information about all data-services participating in DAS.
Based on the information stored in Mapping DB we were able to 
retrieve appropriate data from different data-services, re-map them into 
DAS notations and aggregate them on demand. For instance, the Data 
Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
provide information about CMS blocks. In former case, DBS 
stored information about blocks and their relationships to other 
physics objects, such as dataset and files, while PhEDEx kept
information about its location and replicas.

Upon a user request to 
find a block meta-data, we were able to identify appropriate set of 
DBS and PhEDEx APIs, retrieve data, re-format their output 
into DAS notations and store results into DAS cache. The records
were merged (aggregated) based on identical key-value pairs, e.g.
based on \verb+block.name+ key. Here is an example of such an aggregated record:
\begin{verbatim}
{"das_id": ["4b195404e2194e21c4000030", "4b1953fee2194e21c4000002"], 
 "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
            "subscribed": "n", "custodial": "n", "site": "T1_US_Buffer", 
            "nfiles": "18", "se": "srm.site.com", 
            "size": "112694171838"}, {....}]},
           {"name": "a/b/c#123", "nevents": "570200", 
            "created_by": "USER_DN", "path": "/a/b/c", ...}],
 "das": [{"expire": 1259952908}, {"expire": 1259954702}]}
\end{verbatim}
The \verb+das_id+ refers to api call records stored in Analytics DB,
the \verb+block+ part represents an aggregation block meta-data information from DBS
and PhEDEx systems and the \verb+das+ part contains their expiration time stamps.
Usage of this information was valuable contribution in debugging process of
data-services themselves, e.g. identification of data-service inconsistencies, 
latency studies, etc.

\section{Results\label{Results}}
In this section we discuss preliminary results achieved with DAS system
using the following CMS data-services:
\begin{itemize}
\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
CMS data location and file transfer system;
\item SiteDB \cite{SiteDB}, the CMS site resources;
\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
%\item DQ \cite{DQ}, Data Quality data-service which collects information
%about detector conditions during data taking;
%\item Overview \cite{Overview}, collects information about CMS
%transfer rate, etc.;
\item Dashboard \cite{Dashboard}, the LHC virtual organization
monitoring data-service portal;
\end{itemize}
Each data-service had its own scope, size and peculiar properties. 
For instance, the data in PhEDEx were transient due to constant 
migration of CMS data; the DBS system was divided into dozen 
of individual services (instances) with own size and lifetime 
used by different production teams for various bookkeeping tasks; 
the RunSummary data-service operated on private network accessible 
at CMS online operation center.
%Table \ref{db_size} shows the current size of meta-data from aforementioned 
%data-services.
%\begin{figure}[htb]
%\centering
%\includegraphics[width=100mm]{db_size.pdf}
%\caption{
%Current size of meta-data for CMS data-services participated in DAS, before
%official data-taking. 
%}
%\label{db_size}
%\end{figure}

%\begin{table*}[hbt]
%\centering
%\begin{tabular}{llllll}\hline
%\hline

%Size of  & Dashboard & DBS & PhEDEx & SiteDB & RunSummary \\
%\hline
%tables (GB) & 40.8  & 28.0 & 5.5 & 0.006 & 0.9 \\
%indexes (GB) & 135.9 & 57.2 & 12.6 & 0.004 & 2.7 \\
%\hline
%\hline
%\end{tabular}
%\caption{
%Current size of meta-data for CMS data-services participated in DAS, before
%official data-taking. 
%}
%\label{db_size}
%\end{table*}
At the moment, the amount of meta-data stored in all services is around
75 GB and 200 GB in tables and index sizes, respectively. Although we estimate 
$\sim$500 GB of the meta-data each year during 
normal CMS operation at LHC.
%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
%billion events, 330TB in 32 datasets:
%AlCa/Calibration RAW: 926 million events, 29TB
%Physics RAW: 1.24 billion events, 147TB
%PromptReco: 1.24 billion events, 111TB
%Bulk ALCARECO: 2.34 billion events, 11TB
%Express FEVT: 105 million events, 24TB
%Express ALCARECO: 153 million events, 0.99TB
%HLTMON FEVTHLTALL: 31 million events, 7.4TB
%Above event numbers have significant overlap by design, 
%2.2 billion unique events, including 524 million cosmic muon
%triggers
% 0.3PB - 50GB meta-data in DBS
% 3PB/y - x GB

To test DAS performance we were measured elapsed time and CPU/RAM utilization
required to collect and aggregate information from different
data-services. Table \ref{DAS_benchmark} summarizes the biggest 
contributors so far, the block meta-data information provided by DBS and 
PhEDEx CMS data-services.

\begin{table*}[hbt]
\centering
\begin{tabular}{lllll}\hline
\hline
System & Format & Records & Elapsed time & Elapsed time \\
& & & no cached data & w/ cached data \\
\hline
DAS (DBS) & XML & 382396 & 63 sec & 0.98 sec \\
DAS (PhEDEx) & XML & 186145/180439 & 164 sec & 0.98 sec \\
DAS (aggregate) & JSON & 388108 & 234 sec & 2.05/50.7 sec \\
\hline
\hline
\end{tabular}
\caption{Time required to fetch, parse and aggregate block information
from DBS and PhEDEx systems. In the case of PhEDEx
system we show total number of fetched records 186145 together with
number of records merged in DAS 180439. The total number of DAS records 
were calculated as number of aggregated records plus left overs,
the non-matched records from both systems. The look-up time shown in
last column represents query look-up time (2.05 sec) and time required
to get all records written to disk (50.7 sec).}
\label{DAS_benchmark}
\end{table*}

The initial DAS implementation was done in Python language \cite{Python}
and all tests were performed on x64 Linux platform where
1 CPU was used by DAS server and 1 CPU by MongoDB database server. 
%The CPU utilization
%of DAS server was 67\% in average and RAM usage was below 35MB. 
All tests were performed multiple times to avoid fluctuations. 
The elapsed time measured during those tests consists of the 
following components:
\begin{itemize}
\item[]
{\it
Elapsed time = retrieval time + parsing time + re-mapping time 
        + cache insertion/indexing time 
        + (aggregation time) + (output creation time)
}
\end{itemize}
Here {\it retrieval time} is a time required to access data from data-service,
{\it parsing time} is a time required to read and parse received data, {\it re-mapping time}
is a time required to convert data-service meta-data to DAS records for every object
we parse, {\it cache insertion} and {\it indexing time} represent time spent on DAS 
cache back-end, {\it aggregation time} is required to merge objects into DAS records based
on their common key (\verb+block.name+ in this case) and {\it output creation time}
is a time required to write DAS records to disk. Please note that last two
components, {\it aggregation} and {\it output creation time}, were only applied to
one of the sub-system, either DBS or PhEDEx, depending on insertion order.
% since the first one was injected into
%the fresh database, while the second required the merging step.

@@

The individual tests of DAS components (DBS and PhEDEx) shown roughly 
the same performance. Although the time of PhEDEx component was bigger 
due to longer retrieval and parsing time as well as time spent
during aggregation step.
%The time of PhEDEx component shown in Table \ref{DAS_benchmark} is divided 
%among retrieval and parsing, $\sim$110 sec, and merging steps, $\sim$50 sec,
%respectively. 
As shown in Table \ref{DAS_benchmark},
the look-up time spent to access individual components (last column) was quite
reasonable, roughly 1 second for both systems. While the final time to
get DAS records on disk was about 50 seconds. This was mostly due to I/O and
conversion operations between binary data format used by MongoDB, BSON, to DAS
data format, JSON.

Based on the results of stress tests we measured that we can achieve throughput of
$\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
$\sim$7600 docs/sec for 
reading operations in DAS, which are suitable for our needs. Since all 
operations were performed on a single CPU a further improvements are expected. 
Also, a necessary tuning can be done by using C/C++ driver instead of python one.

\section{Summary}
We presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
at LHC, CERN, Geneva, Switzerland. It was designed to provide caching and
aggregation layer on top of the existing relational and non-relation data-services
mostly in real time fashion. All of the data were retrieved on demand basis,
while data pre-fetching of most common queries is in our plans. We developed
prototype in python for DAS system which is under commissioning phase right now.
The performance studies of DAS system are presented in this paper. 
The CMS experiment is started to collect its data in December of 2009 and 
we expect to accumulate a few PB of real data each year. 
In addition, the Monte-Carlo samples will be produced at this scale.
Based on test studies performed in CMS, we expect that total size of
meta-data produced each year by experiment will be of the order of
500GB. The DAS should sustain such load and will provide generic 
``data-discovery'' service for CMS experiment in years to come.

>>>>

\subsection{DAS Mapping DB\label{MappingDB}}
The Mapping DB is used to track services known to DAS and how DAS should interact with them.
It collects
information about data-service APIs, translates API input/output
parameters into DAS keys and contains mapping between DAS records
and their UI representation.

@@

\subsection{DAS caching system}
The DAS caching system is used to dynamically fetch
and aggregate data upon user requests. We use the following 
algorithm (code snippets are written in Python \cite{Python}):
\begin{enumerate}[1.]
\item Decompose query into set of conditions (\verb+conds+) 
and selection keys (\verb+skeys+).
\item Check cache for query hash. If found, look-up results from the cache,
otherwise proceed to the next step.
\item Identify list of APIs which can serve requested data
\begin{verbatim}
apilist = [r for r in mapping_db(skeys, conds)]
\end{verbatim}
Here \verb+r+ represents a triplet \verb+url, api, args+ and 
\verb+mappind_db+ defines translation between input query and data-service
APIs which can delivery the data.
\item Iterate over APIs
\begin{verbatim}
for url, api, args in apilist:
    if  superset_in_analytics(api, args):
        continue
    for row in parser(get_data(url, api, args)):
        yield row
\end{verbatim}
Here \verb+superset_in_analytics+ checks if provided \verb+api, args+
are subset of already registered in Analytics DB api call(s). The 
\verb+get_data+ and \verb+parser+ are define methods to fetch and parse
data, respectively.
\item Perform aggregation by iterating over records from the previous step
\begin{verbatim}
if cache_has(daskey):
   existing_record = get_from_cache(daskey)
   merge(row, existing_record)
   insert_into_cache(row)
else:
   insert_into_cache(row)
\end{verbatim}
The \verb+daskey+
is a key which identifies the record defined in Mapping DB, e.g. \verb+file.name+
for \verb+file+ record.
We did use a bulk insert operations at this step to increase throughput. 
\item Get results from the cache
\end{enumerate}
%We also have an option to run independent set of daemons, DAS robots, who
%monitor Analytics DB for most popular requests and pre-fetch the most
%common data into the cache. 
We can also automate populating the cache by running an 
independent set of daemons, DAS robots, which monitor the 
Analytics DB for suitable requests (for example queries that are 
popular, periodic in nature or expensive) and pre-fetch data into the cache.

\subsection{DAS queries}
All DAS queries are expressed in a free text-based form, either as a 
set of keywords or key-value pairs, where a pair can represent
condition, e.g. \verb+key > value+. We use a pre-defined set of keys registered 
in DAS to describe common entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. The choice of keys
is based on the analysis of day to day user operations \cite{DBS07}.
Each key represents a DAS record. Every record is stored into the cache
in JSON format. Due to the ``schema-less'' nature of the underlying MongoDB
back-end we are able to store DAS records of arbitrary structure, e.g.
dictionary, lists, key-value pairs, etc. Therefore every DAS key has a 
set of attributes describing its JSON structure. For instance a document
\begin{verbatim}
{"block": {"name": "abc", "size": 1, "replica": [...]}}
\end{verbatim}
is associated with a \verb+block+ key who has the following attributes
\verb+name+, \verb+size+, \verb+replica+. The set of attributes is based 
on a set of meta-data information stored in a record, allowing
for nested structures, e.g. \verb+block.replica.size+. Our users 
are able to specify the context of their queries by using appropriate 
attribute(s). For example, a query for \verb+block+ represents a search for all
block records rather then the ``block'' string, while the query
\verb+block.name=abc*+ instructs the system to look-up appropriate blocks
with matching names. Numerical values, conditions and grouping are
also easy to express, for example, a query
of the form \verb+run > 10 and run < 100+ can be used to
retrieve all run records for the given range of run numbers. 
As a complement to the queries, we will als be adding a dynamic index
of DAS records with a keyword-based search within record content later.

All input queries are transformed into the MongoDB query syntax who is itself
a JSON record. It contains a set of selection 
keys (\verb+fields+) and conditions (\verb+spec+). For example, a query
\verb+dataset, file=abc, run > 10+
is represented as
\begin{verbatim}
{"fields": ["dataset"], "spec": {"file": "abc", "run": {"$gt": 10}}}
\end{verbatim}
Therefore users queries are stored into Analytics DB similar to other DAS 
records and used for further analysis.

\subsection{Data-services and aggregation}
As discussed in section \ref{MappingDB} the DAS Mapping DB 
is the authoritative source of information on the relationships between 
all the data services known to DAS.
%became an authorative
%source of information about all data-services participating in DAS.
Based on the information stored in Mapping DB we are able to 
retrieve appropriate data from different data-services, re-map them into 
DAS notations and aggregate them on demand. For instance, the Data 
Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
provide information about CMS file block structure and locations.
DBS associates groups of files into blocks, while PhEDEx tracks the blocks
location. 
%In the former case, DBS 
%stored information about blocks and their relationships to other 
%physics objects, such as dataset and files, while PhEDEx kept
%information about its location and replicas.

When a query for block meta-data is made, we are
%Upon a user request to 
%find a block meta-data, we were 
able to identify the appropriate set of 
DBS and PhEDEx APIs, retrieve the raw data from the services, format it
for DAS use and add the resulting information to the DAS cache. The records
were merged (aggregated) based on identical key-value pairs, e.g.
based on \verb+block.name+ key. Here is an example of such an aggregated record:
\begin{verbatim}
{"das_id": ["4b195404e2194e21c4000030", "4b1953fee2194e21c4000002"], 
 "block": [{"name": "/a/b/c#123", "replica": [{"complete": "y", 
            "subscribed": "n", "custodial": "n", "site": "T1_US_Buffer", 
            "nfiles": "18", "se": "srm.site.com", 
            "size": "112694171838"}, {....}]},
           {"name": "a/b/c#123", "nevents": "570200", 
            "created_by": "USER_DN", "path": "/a/b/c", ...}],
 "das": [{"expire": 1259952908}, {"expire": 1259954702}]}
\end{verbatim}
The \verb+das_id+ indicates the records stored in the DAS Analytics DB,
the \verb+block+ part represents the aggregated block meta-data information from DBS
and PhEDEx and the \verb+das+ part contains expiration time stamps
about how long the information is valid.
Usage of this information was valuable contribution in debugging process of
data-services themselves, e.g. identification of data-service inconsistencies, 
latency studies, etc.

\section{Results\label{Results}}
In this section we discuss preliminary results achieved with DAS system
using the following CMS data-services:
\begin{itemize}
\item Data Bookkeeping System, DBS \cite{DBS}, the CMS meta-data service;
\item Physics Experiment Data Export, PhEDEx \cite {PhEDEx}, the 
CMS data location and file transfer system;
\item SiteDB \cite{SiteDB}, the CMS site resources;
\item RunSummary \cite{RunSummary}, the CMS run and trigger data-service;
%\item LumiDB \cite{LumiDB} collects information about Luminosity conditions;
%\item DQ \cite{DQ}, Data Quality data-service which collects information
%about detector conditions during data taking;
%\item Overview \cite{Overview}, collects information about CMS
%transfer rate, etc.;
\item Dashboard \cite{Dashboard}, the LHC virtual organization
monitoring data-service portal;
\end{itemize}
Each data-service has its own scope, size and peculiarities. 
For instance, the data in PhEDEx is transient due to constant 
migration of CMS data; the DBS system is divided into several individual 
instances with sizes and lifetimes driven by specific processing tasks; 
the RunSummary data-service is operated on private network accessible 
at CMS online operation center.

At the moment, the amount of meta-data stored in all services is around
75 GB and 200 GB in tables and index sizes, respectively. We estimate to
collect around 500 GB of meta-data each year during 
normal CMS operations. The average query rate to DAS is expected to be at the 
$\leq$10K requests a day.
%In CRAFT09 acquisition era, Tier-0 produced a total of 6.03
%billion events, 330TB in 32 datasets:
%AlCa/Calibration RAW: 926 million events, 29TB
%Physics RAW: 1.24 billion events, 147TB
%PromptReco: 1.24 billion events, 111TB
%Bulk ALCARECO: 2.34 billion events, 11TB
%Express FEVT: 105 million events, 24TB
%Express ALCARECO: 153 million events, 0.99TB
%HLTMON FEVTHLTALL: 31 million events, 7.4TB
%Above event numbers have significant overlap by design, 
%2.2 billion unique events, including 524 million cosmic muon
%triggers
% 0.3PB - 50GB meta-data in DBS
% 3PB/y - x GB

To test DAS performance we measured the elapsed time and CPU/RAM utilization
required to fetch, store and aggregate information from different
data-services. Table \ref{DAS_benchmark} summarizes the biggest 
contributors so far, the block meta-data information provided by DBS and 
PhEDEx CMS data-services.

\begin{table*}[hbt]
\centering
\begin{tabular}{lllll}\hline
\hline
System & Format & Records & Elapsed time & Elapsed time \\
& & & no cached data & w/ cached data \\
\hline
DAS (DBS) & XML & 382396 & 63 sec & 0.98 sec \\
DAS (PhEDEx) & XML & 186145/180439 & 164 sec & 0.98 sec \\
DAS (aggregate) & JSON & 388108 & 234 sec & 2.05/50.7 sec \\
\hline
\hline
\end{tabular}
\caption{Time required to fetch, parse and aggregate block information
from DBS and PhEDEx systems. The PhEDEx
system shows fetched (186145) and merged (180439) records. 
The total number of DAS records are calculated as number of 
aggregated records plus left overs, the non-matched records from both systems. 
The look-up time shown in last column represents query look-up 
time (2.05 sec) and time required to get all records written 
to disk (50.7 sec).}
\label{DAS_benchmark}
\end{table*}

The initial DAS implementation is done in Python language \cite{Python}.
All tests are performed on a 64 bit Linux platform with
1 CPU for the DAS server and 1 CPU for the MongoDB database server. 
All tests are performed multiple times to avoid fluctuations. 
The elapsed time is measured and consists of the following components:
\begin{itemize}
\item[]
{\it
Elapsed time = retrieval time + parsing time + re-mapping time 
        + cache insertion/indexing time 
        + (aggregation time) + (output creation time)
}
\end{itemize}
Here {\it retrieval time} is the time required to access data from data-service,
{\it parsing time} is the time required to read and parse the received data
from the services, and {\it re-mapping time} is the time required to convert 
that information into DAS. {\it Cache insertion} and {\it indexing time} 
represents time spent on DAS back-end caching, {\it aggregation time} is
the time required to merge objects into DAS records based
on their common keys (\verb+block.name+ in this case) and {\it output creation time}
is the time required to write DAS records to disk. Please note that last two
components, {\it aggregation} and {\it output creation time}, are only applied to
one of the sub-system, either DBS or PhEDEx, depending on the insertion order.

@@

The individual tests of DAS components (DBS and PhEDEx) show roughly 
the same performance, although the time taken by the PhEDEx component is longer
due to higher retrieval and parsing time as well as time spent
during aggregation step. As shown in Table \ref{DAS_benchmark},
the look-up time spent to access individual components (last column) is quite
reasonable, roughly 1 second for both systems. While the final time to
get DAS records on disk is about 50 seconds. This is mostly due to I/O and
conversion operations between binary data format used by MongoDB, BSON, to DAS
data format, JSON.

The results of the stress tests have shown that we can achieve a throughput of
$\sim$6000 docs/sec for insertion, $\sim$1100 docs/sec for aggregation and 
$\sim$7600 docs/sec for reading operations in DAS, which is suitable for our 
needs. Please note, we have relatively small audience of users
who perform complex queries, while search engines designed for
a large number of users asking for imprecise, simple data.
Since all tests were performed on a single CPU, further improvements from
expanding to multiple CPUs are expected. In addition, more
performance advantage can be gained by writing dedicated C/C++
python extensions for data intensive operations.

\section{Summary}
%We are presented a new data aggregation service (DAS) developed for CMS High-Energy experiment
%at LHC, CERN, Geneva, Switzerland. It is designed to provide caching and
%aggregation layer on top of the existing relational and non-relation data-services
%mostly in real time fashion. All data are retrieved on demand basis,
%while pre-fetching of most common queries is in our plans. We are developed a
%prototype in Python which is under commissioning phase right now.
%The performance studies of the DAS system are presented in this paper. 
%The CMS experiment is started to collect its data in December of 2009 and 
%we expect to accumulate a few PB of real data each year. 
%In addition, the Monte-Carlo samples will be produced at this scale.
%Based on test studies performed in CMS, we expect that total size of
%meta-data produced each year by experiment will be of the order of
%500GB. The DAS system should sustain such load and provide generic 
%``data-discovery'' service for CMS experiment in years to come.
In this paper we have presented a new data aggregation service (DAS) 
developed for CMS High-Energy experiment at LHC, CERN, Geneva, Switzerland. 
We have developed a prototype for DAS, which has been designed to provide 
caching and aggregation layer on top of the existing relational and 
non-relational data-services in a close to real time fashion. The data 
is retrieved on demand, with pre-fetching of common queries, 
determined from an extensive analytics database, also possible. 
The DAS system is currently being commissioned for production use by CMS.

Data taking at CMS began in December 2009, and, alongside simulation 
data, we expect to accumulate petabytes of data yearly. This will have 
of the order of a terabyte of associated metadata that will need to be 
easily queried by physicists. From the performance studies of DAS presented 
in this paper we expect that the system will be able to sustain such load 
and will provide generic “data-discovery” service for CMS experiment for 
years to come.
